Multilevel selection in causal models: the non-ergodicity of evolutionary and probability theories as the general driver of the emergence of cooperation and specialization.

### 2022-03-09 article.tex
Posterior individuals full model
Typo
Title

%opening
\title{\vspace{-2.2cm}Multilevel selection in causal models: the multiplicative nature of evolutionary and probabilistic selection processes as the general driver for the irreversibility emergence of cooperation and specialization.\footnote{\url{https://github.com/glandfried/transitions/releases/download/doc1.0.1/landfried-\en{transitions}\es{transiciones}.pdf}} }
\author{Gustavo Landfried$^1$}
\affil{\small 1. Universidad de Buenos Aires. Facultad de Ciencias Exactas y Naturales. Departamento de Computaci\'on. Buenos Aires, Argentina}
\affil{\en{Mail}\es{Correo:} \url{glandfried@dc.uba.ar}}

@@

\en{But, what is the characteristic growth rate $g$? }%

@@

\en{The evolutionary problem we are interested in is the selection of cooperative individuals given the environment within each group (level 1), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}, g)$, the selection of groups given the environment (level 2), $P(g^T|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$,  and selection of cooperative individuals given the environment for all groups (multilevel), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$. }%
\es{El problema evolutivo que nos interesa es la selección de los individuos cooperadores dado el ambiente dentro de cada grupo (nivel 1), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}, g)$, la selección de los grupos dado el ambiente (nivel 2), $P(g^T|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$, y la selección de los individuos cooperadores dado el ambiente integrando todos los grupos (multinivel), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$. }%
%
\en{The multilevel selection is obtained by integrating the product of level 1 and 2 selections, }%
\es{La selección multinivel se obtiene integrando el producto de las selecciones de nivel 1 y 2, }%
%
\begin{equation}\label{eq:posterior_multinivel}
\underbrace{P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})}_{\text{\en{Multilevel selection}\es{Selección multinivel}}} = \sum_{g=0}^N \underbrace{P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}, g)}_{\text{\en{Level 1 selection}\es{Selección de nivel 1}}} \cdot \underbrace{P(g^T|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})}_{\text{\en{Level 2 selection}\es{Selección de nivel 2}}}
\end{equation}
%
\en{An alternative way to compute the multilevel selection marginal (Eq.~\ref{eq:posterior_multinivel}) is by integrating the probability of all cooperating individuals given the environments }%
\es{Una forma alternativa de calcular la marginal de la selección multinivel (Eq.~\ref{eq:posterior_multinivel}) es integrando la probabilidad de los individuos cooperadores dado los ambientes }%
%
\begin{equation}\label{eq:posterior_multinivel_alternativa}
\begin{split}
P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}) &= \sum_{j=1}^M P(I^T=j|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})\mathbb{I}(\texttt{coop}(j))
\end{split}
\end{equation}
%
%
\en{where $\mathbb{I}(\cdot)$ is the indicator function. }%
\es{donde $\mathbb{I}(\cdot)$ es la función indiciadora. }%
%
\en{Then, the posteriors of individuals are}\es{Luego} (\en{a detailed mathematical proof can be found in the appendix section ``\nameref{sec:probabilistic_inference}''}\es{una demostración matemática detallada se encuentran en la sección ``\nameref{sec:probabilistic_inference}'' del apéndice.})
%
\begin{equation}
P(k| \vec{\Aa}^{\,1}, \dots, \vec{\Aa}^T) = 
\begin{cases}
P(k)\prod_{t=1}^{T} P(k|\vec{\Aa}^{\,t}) &  r=0  \\
P(k)\prod_{t=1}^{T} \sum_j^{\texttt{\en{partners}\es{socios}}(r)} \frac{1}{N} P(j|\vec{\Aa}^{\,t}) & r  = N  \\
\Big(P(k)\prod_{t=1}^{T} P(k|\vec{\Aa}^{\,t}) \Big) + \Big(\sum_{t=1}^{T} P(c|\wedge_{q=1}^t\vec{\Aa}^{\,q})  \prod_{q=t+1}^T P(k|\vec{\Aa}^{\,q}) \Big) & 0 < r < N  
\end{cases}
\end{equation}
%
\en{where $r = \texttt{region}(k)$ is the region to which individual $k$ belongs, $\texttt{partners}(r)$ is the set of cooperating individuals in region $r$, and $c$ is a cooperative individual belonging to that specific region, $c \in \texttt{partners}(r)$. }%s
\es{donde $r = \texttt{region}(k)$ es la región a la que pertenece el individuo $k$, $\texttt{\en{partners}\es{socios}}(r)$ es el conjunto de individuos cooperadores de la región $r$, y $c$ es un individuo cooperador de esa región, $c \in \texttt{\en{partners}\es{socios}}(r)$. }%

>>>>

%opening
\title{\vspace{-2.2cm}Multilevel selection in causal models: the multiplicative nature of evolutionary and probabilistic selection processes as the general driver for the irreversible emergence of cooperation and specialization.\footnote{\url{https://github.com/glandfried/transitions/releases/download/doc1.0.1/landfried-\en{transitions}\es{transiciones}.pdf}} }
\author{Gustavo Landfried$^1$}
\affil{\small 1. Universidad de Buenos Aires. Facultad de Ciencias Exactas y Naturales. Departamento de Computaci\'on. Buenos Aires, Argentina}
\affil{\en{Mail}\es{Correo:} \url{glandfried@dc.uba.ar}}

@@

\en{But, which is the characteristic growth rate $g$? }%

@@

\en{The evolutionary problem we are interested in is the selection of cooperative individuals given the environment within each group (level 1), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}, g)$, the selection of groups given the environment (level 2), $P(g^T|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$,  and selection of cooperative individuals given the environment for all groups (multilevel), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$. }%
\es{El problema evolutivo que nos interesa es la selección de los individuos cooperadores dado el ambiente dentro de cada grupo (nivel 1), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}, g)$, la selección de los grupos dado el ambiente (nivel 2), $P(g^T|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$, y la selección de los individuos cooperadores dado el ambiente integrando todos los grupos (multinivel), $P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})$. }%
%
\en{The multilevel selection is obtained by integrating the product of level 1 and 2 selections, }%
\es{La selección multinivel se obtiene integrando el producto de las selecciones de nivel 1 y 2, }%
%
\begin{equation}\label{eq:posterior_multinivel}
\underbrace{P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})}_{\text{\en{Multilevel selection}\es{Selección multinivel}}} = \sum_{g=0}^N \underbrace{P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}, g)}_{\text{\en{Level 1 selection}\es{Selección de nivel 1}}} \cdot \underbrace{P(g^T|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})}_{\text{\en{Level 2 selection}\es{Selección de nivel 2}}}
\end{equation}
%
\en{An alternative way to compute the multilevel selection marginal (Eq.~\ref{eq:posterior_multinivel}) is by integrating the probability of all cooperating individuals given the environments }%
\es{Una forma alternativa de calcular la marginal de la selección multinivel (Eq.~\ref{eq:posterior_multinivel}) es integrando la probabilidad de los individuos cooperadores dado los ambientes }%
%
\begin{equation}\label{eq:posterior_multinivel_alternativa}
\begin{split}
P(\texttt{coop}(i^T)|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1}) &= \sum_{j=1}^M P(I^T=j|\vec{\Aa}^{\,1}, \dots, \vec{\Aa}^{\,T-1})\mathbb{I}(\texttt{coop}(j))
\end{split}
\end{equation}
%
%
\en{where $\mathbb{I}(\cdot)$ is the indicator function. }%
\es{donde $\mathbb{I}(\cdot)$ es la función indiciadora. }%
%
\en{Then, the posteriors of individuals are}\es{Luego} (\en{a detailed mathematical proof can be found in the appendix section ``\nameref{sec:probabilistic_inference}''}\es{una demostración matemática detallada se encuentran en la sección ``\nameref{sec:probabilistic_inference}'' del apéndice.})
%
\begin{equation}
P(k| \vec{\Aa}^{\,1}, \dots, \vec{\Aa}^T) = 
\begin{cases}
%P(k)\prod_{t=1}^{T} P(k|\vec{\Aa}^{\,t}) &  r=0  \\
P(k)\prod_{t=1}^{T} \sum_j^{\texttt{\en{partners}\es{socios}}(r)} \frac{1}{N} P(j|\vec{\Aa}^{\,t}) & \texttt{coop}(k)  \\ % r = N
\Big(P(k)\prod_{t=1}^{T} P(k|\vec{\Aa}^{\,t}) \Big) + \Big(\sum_{t=1}^{T} P(c|\wedge_{q=1}^t\vec{\Aa}^{\,q})  \prod_{q=t+1}^T P(k|\vec{\Aa}^{\,q}) \Big) & \neg\texttt{coop}(k) % 0 < r < N  
\end{cases}
\end{equation}
%
\en{where $r = \texttt{region}(k)$ is the region to which individual $k$ belongs, $\texttt{partners}(r)$ is the set of cooperating individuals in region $r$, and $c$ is a cooperative individual belonging to that specific region, $c \in \texttt{partners}(r)$. }%s
\es{donde $r = \texttt{region}(k)$ es la región a la que pertenece el individuo $k$, $\texttt{\en{partners}\es{socios}}(r)$ es el conjunto de individuos cooperadores de la región $r$, y $c$ es un individuo cooperador de esa región, $c \in \texttt{\en{partners}\es{socios}}(r)$. }%
